Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-11-30 Response_referees.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

List of major changes:
\begin{itemize}
 \item We have further clarified the relation between localization properties and features of the computational basis needed
 to provide precise results. Major modifications under this perspective have been done in the introduction, in section II at pag. 5 before the analysis of fig. 2,in section IV and in the conclusions.
 \item The introduction has been deeply revised. In particular we have reduced the length of the first two paragraph concerning the differences between accuracy and precision.
 \item We added a comment after eq. (7) on the non-hermiticity of the response density and corrected the equation (11).
 \item Added a dedicated subsection (actually the IIA at page 4)  to discuss the localization properties of the FS states for static perturbations.
 \item Added a panel in figure 2 that contains the plot of the real part of $\alpha(\omega)$ in calculated in different computational domains. This new plot aims to evidence that the localization
 properties of th FS affect, on equal footing, both the real and the imaginary part of the linear response functional.
 \item We have reorganized the material after section III A. In particular we have introduced a new section (number IV in the actual version of the paper) in which we have discussed both the physical features of the excitation spectrum and the relative importance of localized vs. delocalized sector in expressing the susceptibility functional. This section also collects computational results for molecules beyond the case of $CO$ and benzene.
 \item We have also included in our analysis the treatment of various molecular systems, of various sizes and symmetries, such as to further illustrate the generality of our assumptions. A table has been added for comparing the excitation landscape of the molecules, in particular related to the ability of the numerical excitations to
 express the static polarizability tensor.
\end{itemize}

@@

\section*{Response to the first referee}

We wish to thank the referee for the positive review on the paper. The text has been deeply revised according to the suggestions of both the referees.

Here we report the detailed answers concerning the main aspects raised in the referee's report:
\begin{enumerate}
  \item \emph{on the point concerning the dimension of the simulation box needed to achieve the convergence of the delocalized states of the $CO$ molecule}.

  Delocalized fluctuation states do not achieve convergence as long as the simulation box is increased. Since they behave as continuum states they suffer of the same continuum collapse behavior described in section IA for the unbound solution of the Schr\"odinger equation. What is expected to converge is the response functional, in the particular case presented here the real and imaginary part of the dynamical polarizability.
  The value of convergence is system/observable and perturbation dependent.
  % It can be found through an explicit convergence analysis by introducing an acceptance tolerance, however this procedure is out of the scope of our analysis and has not been performed.

  Nonetheless, to have an idea of the dimension of the box needed to achieve convergence we can refer to the case of the Benzene molecule and make usage
  of the result found in our analysis, compared to the one reported in fig. 2 of \cite{baroni2008}. Within our analysis we show that a box linear size of about 18 Angostroem is sufficient to produce converged results for excitation spectrum below IP (around 6.5 eV in the present case). On the contrary, results of \cite{baroni2008} evidence that to achieve convergence up to 30 eV a box of linear dimension of more than 30 Angstroem is needed, so going from the localized to the delocalized sector of excitation spectrum requires a relevant increase in the simulation domain.

@@

We answer to the critical points raised out in the report of the referee:

\begin{enumerate}

 \item \emph{On the localized sector of excitation. Dependence on the symmetry and dimension of the molecule.}

 % In section IIIA we have provided formal arguments that evidence that the excitation spectrum splits into two sectors with distinct features. The first one
 % is composed by a finite set of localized excitations with discrete eigenvalues, whereas the second one contains a continuum spectrum of delocalized objects.
 % For excitation energies lower than the ionization potential (IP) of the molecule all the excitations belong to the first sector. In the intermediate energy range which goes from IP to the absolute value of the energy of the deepest occupied orbital the two sectors compenetrate so that there are localized excitation embedded in a continuum of delocalized ones, and finally only elements of the second sector exist for higher values of the energy.
 %
 % These formal results have been supported by numerical calculations in which the excitation spectrum of several molecules has been explicitly computed.
 We have added in the Appendix the explaination
 of the procedure we employed for the determination of the excitation landscape.
  In the previous version of the paper this type of analysis was limited to the case of CO and Benzene molecules.

  However the same structure of the excitation spectrum is quite general and turns out not to be affected by the size or symmetry of the molecule. To check this statement we have repeated
 the same computational analysis for several other molecules of different shape and dimension like the Fullerene and the Aflatoxin, a low-symmetry molecule of 35 atoms which is not as little and exhibit generic features. Results for this studies
 are presented as supplementary materials. See Ref.XXX

 Looking at the excitation landscape for these cases, extracted with the same procedure and tolerance parameters of the cases presented in the paper we see that our consideration are relatively independent from the molecule symmetry: in particular, the number of excitations which
 belong to the localized sector, even above the first IP
 seems to _increase_ with the molecule's size.
 This is due to the increasing of the KS density of states below the continuum threshold which ensures the realization of more localized excitations.

However, see answer below, such a behaviour is not particularly relevant as we have demonstratated that the
localized sector alone is --- in most of the cases of interest --- not sufficient to represent the properties of the linear susceptibility.

@@

\item \emph{Regarding the localized excitations which can still be found above the IP threshold (Fig 5a), is there any scope to target these, to the
exclusion of other states, so that a set of just the localized excitations can be calculated at reduced computational cost compared to the whole set?}

In the previous version of the manuscript the answer to this question was adressed in the last section before the conclusions.

 During the review process we have deeply revised the final part of the paper (in particular the section IVB in the actual version of the manuscript) in order to further  clarify which is the relevance the two excitation sectors for obtaining an effective representation of the susceptibility functional.
 For a sake of concreteness this analysis is declined for the relevant case of the computation of the dynamical polarizability.

 It emerges that the localized
 sector of excitation is sufficient to express the imaginary part of $\alpha(\omega)$ \emph{only} for values of $\omega$ below IP (see eq. 21). The evaluation of
 other quantities, namely $\mathrm{Im}(\alpha(\omega))$ above IP or $\mathrm{Re}(\alpha(\omega))$ in all the $\omega$ range requires a relevant contribution from
 both the sectors of excitation spectrum. So, to answer to the question raised by the referee we can state that localized excitations above IP can be isolated and computed using the procedure described in the previous point of this report, but their knowledge alone is only useful for the extraction of absorption spectra in the optical regime.

@@

 \item \emph{On the exclusion ''by design'' of the delocalized part of the excitation spectrum}

 For this reason, we have further clarified the
 importance of a computational setup that does not exclude ``by design'' the long-range oscillatory behaviour of the localized excitations, especially for the representation
 of the fluctuation states above threshold.
 Nonetheless, in the optical regime the FS are localized quantities, and it is thus possible -in principle - to express them precisely, in localized basis sets. However, such a representation is \emph{doomed to fail} for frequencies above IP, due to the incorrect asymptotic behaviour of  states above threshold.

### 2018-12-03 AnalyticStructure_main.tex
response to referees drafted, only the git repo has to be created
revision

We believe that these considerations, based on simple manipulations on the fundamental LR equations, will help in increase the reliability of LR
calculations in the community, and in building test-sets that would help in the \emph{calibration} of present and future computer codes, thereby
increasing the predictivity of present-day theoretical approaches. The example of the static polarizability of molecules
presents ideal features as an initial playground. Work is ongoing in this direction.

@@

>>>>

We believe that these considerations, based on simple manipulations on the fundamental LR equations, will help in increase the reliability of LR
calculations in the community, and in building test-sets that would help in the \emph{calibration} of present and future computer codes, thereby
increasing the predictivity of present-day theoretical approaches. The example of the static polarizability of molecules
presents ideal features as an initial playground, as it has already been pointed out recently~\cite{HeadGordon}. Work is ongoing in this direction.

@@

All the details of the calculations, the \textsl{Jupyter} notebooks of the data analysis, including the supplementary informations, may be found at the URL~\footnote{\texttt{https://github.com/luigigenovese/LR-nb/blob/master/ POLARIZABILITY/nb\_paper/LR\_Analysis.ipynb}, or \texttt{https://gitlab.com/luigigenovese/ localization-of-excitations/ blob/master/LR\_Analysis.ipynb}}
